Laplacian envelope

### 2021-11-23 Notes/laplace.tex
typos

Note that, for any state \(i \in C_a\)
%
\begin{equation*}
\begin{aligned}
  [\W\pot \etw]_i &= \sum_{j \in C_{a+1}} \Wm\pot_{ij} (\etwm_j - \etwm_i)
    = (\etwm_{a+1} - \etwm_a) \sum_{\mathclap{j \in C_{a+1}}} \Wm\pot_{ij}
    = (\etwm_{a+1} - \etwm_a) \lambda\pot_i, \\
  [\W\dep \etw]_i &= (\etwm_{a-1} - \etwm_a) \lambda\dep_i,
  \quad\text{where} \quad
  \lambda^\nu_i =  \sum_{j \neq i} \Wm^\nu_{ij}.
\end{aligned}
\end{equation*}
%
which also applies wth \(\etw\) replaced with \(\thb\) or \(\mathbf{c}\).
All models, not just the optimal ones, satisfy
%
\begin{equation*}
\begin{aligned}
  \frg \thb(0) &= (\I - \onev \eq) \mathbf{a} = - \enc \etw(0) - \CA,
  \\
  \frg \etw(0) &= (\I - \onev \eq) \w = \w - \Delta p\, \onev.
\end{aligned}
\end{equation*}
%
Using the results above
%
\begin{equation}\label{eq:dgn_lump}
\begin{gathered}
\begin{aligned}
  (\thbm_{a+1} - \thbm_a) \lambda\pot_i + (\thbm_{a-1} - \thbm_a) \lambda\dep_i
    &= - (\etwm_{a+1} - \etwm_a) \lambda\pot_i + (\etwm_{a-1} - \etwm_a) \lambda\dep_i - \CA,
  \\
  (\etwm_{a+1} - \etwm_a) \lambda\pot_i + (\etwm_{a-1} - \etwm_a) \lambda\dep_i &= \wm_a - \Delta p,
\end{aligned}
\\
  \implies
\begin{aligned}
  \lambda\pot_i &= \frac{(\CA - \wm_a + \Delta p) (\etwm_a - \etwm_{a-1}) 
                        + (\wm_a - \Delta p) (\thbm_a - \thbm_{a-1})}
                      {(\etwm_{a+1} - \etwm_a) (\thbm_a - \thbm_{a-1})
                        - (\etwm_a - \etwm_{a-1}) (\thbm_{a+1} - \thbm_a)
                        - 2 (\etwm_{a+1} - \etwm_a) (\etwm_a - \etwm_{a-1})}, \\
  \lambda\dep_i &= \frac{(\CA + \wm_a - \Delta p) (\etwm_{a+1} - \etwm_a) 
                        + (\wm_a - \Delta p) (\thbm_{a+1} - \thbm_a)}
                      {(\etwm_{a+1} - \etwm_a) (\thbm_a - \thbm_{a-1})
                        - (\etwm_a - \etwm_{a-1}) (\thbm_{a+1} - \thbm_a)
                        - 2 (\etwm_{a+1} - \etwm_a) (\etwm_a - \etwm_{a-1})}.
\end{aligned}
\end{gathered}
\end{equation}
%
Because the right-hand sides above depend on \(a\) and not \(i\), \cref{eq:dgn_lump} implies that \(\boldsymbol{\lambda}^\nu\) also has the same degeneracy as \(\etw\), which is the condition for lumpability under the partitions \(\brc{C_a}\) \cite{kemeny1960finite,Ball1993Lumpability,burke1958markovian}.

@@

First let us maximise \cref{eq:serialarea} at fixed $\eqm_\pm = \sum_k \eqm_k \prn{\frac{1\pm\wm_k}{2}}$.
Clearly this will happen when we put all of the probability at the ends: $\eqm_1=\eqm_-$ and $\eqm_n=\eqm_+$ are the only non-zero $\eqm_k$.
This gives an area of
\footnote{Note that including the dropped factors in \cref{eq:lsnr} would only change this to \(\CA \leq (M-1)(4\eqm_+\eqm_-)/r\sqrt{1 - (\Delta f)^2(\eqm_+ - \eqm_-)^2}\), whose maximum also lies at \(\eqm_+=\eqm_-= \frac{1}{2}\).}
%
\begin{equation}\label{eq:serialextarea}
  \CA \leq (M-1) \prn{4\eqm_+\eqm_-} = (M-1)\prn{1 - (\eqm_+ - \eqm_-)^2}.
\end{equation}
%
This is maximised at \(\eqm_+=\eqm_-= \frac{1}{2}\):
%
\begin{equation}\label{eq:maxarea}
  \CA \leq (M-1).
\end{equation}
%
Note that the markov process that achieves this is not ergodic, the two states at each end are absorbing. 
This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.

>>>>

Note that, for any state \(i \in C_a\)
%
\begin{equation*}
\begin{aligned}
  [\W\pot \etw]_i &= \sum_{\mathclap{j \in C_{a+1}}} \Wm\pot_{ij} (\etwm_j - \etwm_i)
    = (\etwm_{a+1} - \etwm_a) \sum_{\mathclap{j \in C_{a+1}}} \Wm\pot_{ij}
    = (\etwm_{a+1} - \etwm_a) \lambda\pot_i, \\
  [\W\dep \etw]_i &= (\etwm_{a-1} - \etwm_a) \lambda\dep_i,
  \quad\text{where} \quad
  \lambda^\nu_i =  \sum_{j \neq i} \Wm^\nu_{ij}.
\end{aligned}
\end{equation*}
%
which also applies wth \(\etw\) replaced with \(\thb\) or \(\mathbf{c}\).
All models, not just the optimal ones, satisfy
%
\begin{equation*}
\begin{aligned}
  \frg \thb(0) &= (\I - \onev \eq) \mathbf{a} = - \enc \etw(0) - \CA,
  \\
  \frg \etw(0) &= (\I - \onev \eq) \w = \w - \Delta p\, \onev.
\end{aligned}
\end{equation*}
%
Using the results above
%
\begin{equation}\label{eq:dgn_lump}
\begin{gathered}
\begin{aligned}
  (\thbm_{a+1} - \thbm_a) \lambda\pot_i + (\thbm_{a-1} - \thbm_a) \lambda\dep_i
    &= - (\etwm_{a+1} - \etwm_a) \lambda\pot_i + (\etwm_{a-1} - \etwm_a) \lambda\dep_i - \CA,
  \\
  (\etwm_{a+1} - \etwm_a) \lambda\pot_i + (\etwm_{a-1} - \etwm_a) \lambda\dep_i &= \wm_a - \Delta p,
\end{aligned}
\\
  \implies
\begin{aligned}
  \lambda\pot_i &= \frac{(\CA - \wm_a + \Delta p) (\etwm_a - \etwm_{a-1}) 
                        + (\wm_a - \Delta p) (\thbm_a - \thbm_{a-1})}
                      {(\etwm_{a+1} - \etwm_a) (\thbm_a - \thbm_{a-1})
                        - (\etwm_a - \etwm_{a-1}) (\thbm_{a+1} - \thbm_a)
                        - 2 (\etwm_{a+1} - \etwm_a) (\etwm_a - \etwm_{a-1})}, \\
  \lambda\dep_i &= \frac{(\CA + \wm_a - \Delta p) (\etwm_{a+1} - \etwm_a) 
                        + (\wm_a - \Delta p) (\thbm_{a+1} - \thbm_a)}
                      {(\etwm_{a+1} - \etwm_a) (\thbm_a - \thbm_{a-1})
                        - (\etwm_a - \etwm_{a-1}) (\thbm_{a+1} - \thbm_a)
                        - 2 (\etwm_{a+1} - \etwm_a) (\etwm_a - \etwm_{a-1})}.
\end{aligned}
\end{gathered}
\end{equation}
%
Because the right-hand sides above depend on \(a\) and not \(i\), \cref{eq:dgn_lump} implies that \(\boldsymbol{\lambda}^\nu\) also has the same degeneracy as \(\etw\), which is the condition for lumpability under the partitions \(\brc{C_a}\) \cite{kemeny1960finite,Ball1993Lumpability,burke1958markovian}:
for any two states in \(C_a\), we get the same value for the sum of transition rates from that state to all states in \(C_b\).

@@

First let us maximise \cref{eq:serialarea} at fixed $\eqm_\pm = \sum_k \eqm_k \prn{\frac{1\pm\wm_k}{2}}$.
Clearly this will happen when we put all of the probability at the ends: $\eqm_1=\eqm_-$ and $\eqm_M=\eqm_+$ are the only non-zero $\eqm_k$.
This gives an area of
\footnote{Note that including the dropped factors in \cref{eq:lsnr} would only change this to \(\CA \leq (M-1)(4\eqm_+\eqm_-)/r\sqrt{1 - (\Delta f)^2(\eqm_+ - \eqm_-)^2}\), whose maximum also lies at \(\eqm_+=\eqm_-= \frac{1}{2}\).}
%
\begin{equation}\label{eq:serialextarea}
  \CA \leq (M-1) \prn{4\eqm_+\eqm_-} = (M-1)\prn{1 - (\eqm_+ - \eqm_-)^2}.
\end{equation}
%
This is maximised at \(\eqm_+=\eqm_-= \frac{1}{2}\):
%
\begin{equation}\label{eq:maxarea}
  \CA \leq (M-1).
\end{equation}
%
Note that the markov process that achieves this is not ergodic, the two states at each end are absorbing. 
This is similar to the results found numerically in \cite{Barrett2008discrete} in a slightly different situation.
